---
title: "Project"
author: "Chun Kang Lu 1008161150"
date: "2023-03-24"
output: html_document
---

```{r, echo=F}
library(tidyverse)
```

```{r}
# We have to note the data was downloaded on March 24, and the dataset is updated every week, so future downloads may get slightly different results.
data = read.csv("COVID19 cases.csv")
head(data)
```
```{r}
glimpse(data)
```
```{r}
# Turn Episode.Date and Reported.Date into Date type
data = data %>% mutate(Episode.Date = as.Date(Episode.Date, format = "%Y-%m-%d"),
                       Reported.Date = as.Date(Reported.Date, format = "%Y-%m-%d"))
head(data)
```
```{r}
dim(data)
```

# About the Dataset

Available [here](https://open.toronto.ca/dataset/covid-19-cases-in-toronto/), the COVID-19 Cases dataset provides weekly updating information about all reported cases of COVID-19 by the Toronto Public Health since the first reported case in January of 2020. The information in this dataset changes every week as new information get updated at 8:30 AM every Tuesday. For consistency in our analysis, we will be using the downloaded csv version of this dataset from March 24 in this report.

This dataset contains 15 columns and 394761 rows, with each row representing a reported individual with COVID-19.

The columns are as follows:
* `_id` - A unique row identifier for Open Data database.
* `Assigned_ID` - A unique ID assigned to cases by Toronto Public Health for the purposes of posting to Open Data, to allow for tracking of specific cases.
* `Outbreak Associated` - Cases associated with outbreaks are identified by their settings and institutions (eg. long-term care homes, retirement homes, hospitals, homeless shelters, etc.).
* `Age Group` - Age at time of illness. Age groups (in years): ≤19, 20-29, 30-39, 40-49, 50-59, 60-69, 70-79, 80-89, 90+, unknown (blank)
* `Neighbourhood Name` - Toronto is divided into 140 geographically distinct neighborhoods that were established to help government and community agencies with local planning by providing socio-economic data for a meaningful geographic area.
* `FSA` - Forward sortation area (i.e. first three characters of postal code) based on the case's primary home address.
* `Source of Infection` - The most likely way the person contracted COVID-19. These are determined through a few factors of a public health investigator's assessment, association with a confirmed COVID-19 outbreak, and known risk factors (like travel or close contact with another known COVID-19 case).
* `Classification` - Categorization of the cases as confirmed or probable, according to standard criteria of the Ontario Ministry of Health.
* `Episode Date` - Best estimate to when the disease was first acquired.
* `Reported Date` - Date which the case was reported to Toronto Public Health.
* `Client Gender` - Self-reported gender of the individual.
* `Outcome` - 3 scenarios of fatal, resolved, and active. 
    * Fatal means the individual has died directly due to COVID-19 and not an alternative cause of death. 
    * Resolved refers to individuals who have either:
        * Been marked Fatal but the cause of death was not due to COVID-19.
        * Are alive and has been more than 14 days from the episode date and the case is not currently hospitalized, intubated, or in ICU.
    * Active refers to all other ongoing cases
* `Ever Hospitalized` - Cases that were hospitalized related to their COVID-19 infection (includes cases that are currently hospitalized and those that have been discharged or are deceased).
* `Ever in ICU` - Cases that were admitted to the intensive care unit (ICU) related to their COVID-19 infection (includes cases that are currently in ICU and those that have been discharged or are deceased).
* `Ever Intubated` - Cases that were intubated related to their COVID-19 infection (includes cases that are currently intubated and those that have been discharged or deceased)


